Towards a Machine Learning approach for Aggregate Computing

### 2021-11-27 paper-2021-woa-ac-ml.tex
add future works and discussion

\section{Introduction}
Recent trends like pervasive~\cite{DBLP:journals/wc/Satyanarayanan01} and autonomic~\cite{DBLP:journals/computer/KephartC03} computing 
 promote a vision of artificially designed Collective Adaptive Systems (CASs),
 namely collections of agents coordinating and adapting %as wholes 
 to pursue joint goals in changing environments.
% due to the increasing interconnected computational devices placed in ever-changing environments.
 Examples of CASs include robot swarms, smart cities, and digitally-augmented crowds.
%
%CASs are characterised by 
%\begin{itemize}
%\item a global order that emerges from local node interactions, 
%\item  decentralised control, and 
%\item adaptive behaviours towards environmental changes to pursuit a collective goal.
%\end{itemize}
%
CAS are notoriously hard to design and control 
 %due to as part of complex systems
 as globally desired behaviour
 has to emerge from local behaviours and interactions. 
For this reason, over the years, various approaches have tried to handle this complexity. Some of them take inspiration from the natural system composed of social animals --- like bees, fish, and ants. 
%
In this case, a sort of hive mind arises -- 
 also referred to as swarm intelligence --  
 born from local interactions between agents. 
%
These kinds of behaviours are accomplished with self-organisation, 
 namely a global order that emerges from continuous 
 interactions between simple entities.

@@

Collective Adaptive Systems~\cite{DBLP:conf/huc/Ferscha15} are characterized by a very high number of nodes without a central coordinating authority that coordinate nodes. The overall behaviour is a consequence of local interactions. The systems expose self-* properties.

@@

In this work, we focus on Aggregate Computing~\cite{DBLP:journals/computer/BealPV15}, a global-to-local approach where
 the self-organising collective behaviour is expressed in terms of manipulation of \textit{computational fields}.
The latter could be seen as the extension of electromagnetic fields to Computer Science since it consists of a 
 map from any devices to their computational value (i.e. the result of a computation).
The computational field manipulation is ruled by \textit{field calculus}, which define the constructs needed 
 to express spatio-temporal computation~\cite{DBLP:conf/coordination/AudritoBDV18}.
The novelty of AC consists in its composable (functional-like) system-size independent program definition: 
using basic constructs we identify the main patterns (\textit{building-blocks}) that could be reusable at the application level.
By opportunistically crafting these building-blocks it is possible to verify prominent properties such as eventual consistency~\cite{DBLP:journals/taas/BealVPD17} 
 (i.e. the ensemble behaviour is independent of the underlying network details)
 and self-stabilisation~\cite{DBLP:journals/tomacs/ViroliABDP18} (i.e. the ability of a system to recover from arbitrary changes).

@@

\section{Discussion and Concluding Remarks}\label{conclusion}
In this article, we have discussed a preliminary example of how AC can be integrated with RL. 
We have then compared AC problems with current Machine Learning solutions in MASs systems.
%
We have defined a possible roadmap towards full support of learning mechanisms in AC. %
As a proof of concept, we have discussed preliminary use case where 
 learning successfully enhances a standard AC program.
Furthermore, we have shown that the learnt program can easily scale up to system size.

%\subsection{Roadmap}
%The path towards full integration of Machine Learning and Aggregate Computing is long and probably we do not know the implications that it can bring. 
%Currently, we program to:
%\begin{enumerate}
%  \item Enhance building blocks through learning -- \textit{offline}: 
%  in this first step, we need to understand what kind of learning modality is most suitable for us. 
%  To decrease the complexity, we aim at perform learning \textit{offline} and at the \textit{building block} level.
%  \item Enhance building blocks through learning -- \textit{online}:
%  here, the most challenging part in bringing the offline results could be in the reward function definition.
%  Indeed, at runtime, we cannot use a global view of the system. 
%  A novel way to describe a collective reward by Aggregate Computing might be taken into consideration.
%  \item Create building blocks through learning: some building blocks currently lack a proper implementation -- 
%  such as the distributed leader election -- so learning can help us also in crafting them. 
%  Here, though, the research space increase drastically, 
%  so probably it might be not easy to bring our previous results there.
%  \item Improve framework aspects through learning: 
%  when we successfully apply to learn in the first two steps, 
%  we can also consider moving learning at the framework level, to manage non-functional aspects. 
%  In this way, developers can focus on the "high level" behaviour aspect.
%\end{enumerate}

\todos{discussion about the result}
\todos{limitations: it does not easily scale with the complex application, it cannot be applied online, it cannot be used in different scenarios}
\todos{future work: integrate deep learning solution, integrate actor-critic method (team learning)}
\todos{future work: move learning at the framework level}

%%
%% Define the bibliography file to be used

>>>>

\section{Introduction}
Recent trends like pervasive~\cite{DBLP:journals/wc/Satyanarayanan01} and autonomic~\cite{DBLP:journals/computer/KephartC03} computing 
 promote a vision of artificially designed Collective Adaptive Systems (CASs)~\cite{DBLP:conf/huc/Ferscha15},
 namely collections of agents coordinating and adapting %as wholes 
 to pursue joint goals in changing environments.
% due to the increasing interconnected computational devices placed in ever-changing environments.
 Examples of CASs include robot swarms, smart cities, and digitally-augmented crowds.
%
%CASs are characterised by 
%\begin{itemize}
%\item a global order that emerges from local node interactions, 
%\item  decentralised control, and 
%\item adaptive behaviours towards environmental changes to pursuit a collective goal.
%\end{itemize}
%
CAS are notoriously hard to design and control 
 %due to as part of complex systems
 as globally desired behaviour
 has to emerge from local behaviours and interactions. 
For this reason, over the years, various approaches have tried to handle this complexity. Some of them take inspiration from the natural system composed of social animals --- like bees, fish, and ants. 
%
In this case, a sort of hive mind arises -- 
 also referred to as swarm intelligence --  
 born from local interactions between agents. 
%
These kinds of behaviours are accomplished with self-organisation, 
 namely a global order that emerges from continuous 
 interactions between simple entities.

@@

Collective Adaptive Systems are characterized by a very high number of nodes without a central coordinating authority that coordinate nodes. The overall behaviour is a consequence of local interactions. The systems expose self-* properties.

@@

In this work, we focus on Aggregate Computing~\cite{DBLP:journals/computer/BealPV15}, a global-to-local approach where
 the self-organising collective behaviour is expressed in terms of manipulation of \textit{computational fields}.
The latter could be seen as the extension of electromagnetic fields to Computer Science since it consists of a 
 map from any devices to their computational value (i.e. the result of a computation).
The computational field manipulation is ruled by \textit{field calculus}, which define the constructs needed 
 to express spatio-temporal computations~\cite{DBLP:conf/coordination/AudritoBDV18}.
The novelty of AC consists in its composable (functional-like) system-size independent program definition: 
using basic constructs we identify the main patterns (\textit{building-blocks}) that could be reusable at the application level.
By opportunistically crafting these building-blocks it is possible to verify prominent properties such as eventual consistency~\cite{DBLP:journals/taas/BealVPD17} 
 (i.e. the ensemble behaviour is independent of the underlying network details)
 and self-stabilisation~\cite{DBLP:journals/tomacs/ViroliABDP18} (i.e. the ability of a system to recover from arbitrary changes).

@@

\section{Discussion and Concluding Remarks}\label{conclusion}
In this article, we have discussed the combination of AC with Machine Learning techniques in order to improve adaptivity
 and effictveness of the AC framework.
In doing so, we compared AC problems with current Machine Learning solutions in MASs systems.
%
Finally, we explore the combination of AC with RL (in particular Hysteretic Q-Learning).
RL application in this kind of settings is quite an explored -- even if some works exists~\cite{DBLP:conf/icml/YangLLZZW18} -- 
 area for several reason, such us the highly system dynamic, non-stationarity,  partial enviroment observability, and non-bounded system size.
In a case study we empirically shown the effictveness of RL, by which we speed up the convergence time of
 aggregate program taking in consideration -- the hop count.
Furthermore, we have shown that the learnt program can easily scale up to system size.

This is a first effort towards an ``hybrid'' collective programming approach, where the high-level program specification is 
 done by a designer and the low-level matters are leaved to machine learning algorithms.

We know that this is preliminary work and it shed light on several problems that we should handle in the future.

Aggregate systems are partially observable from the perspective of each node. 
 Local observations may not be statistically significant in summarising the state of the entire aggregate. 
 In the future, we could address this problem through recurrent networks and a CTDE learning methodology.

The learning, in this case, was applied at the building block level. 
 Another possible direction is to use it at the framework level, e.g., to select which messages to send to the neighbourhood or not. 
 This has already been addressed in the literature, usually under the name of \textit{learn to communicate}~\cite{DBLP:journals/corr/abs-1908-03963}. 
 A novel way to handle this problem is to use Graph Neural Networks~\cite{DBLP:journals/corr/abs-1812-08434} to learn the representation of a distributed system and to select messages based on that representation. 
 Thus, we should consider this methodology when we will extend RL at the AC framework level.

In this case, we applied a variation of the independent learning approach. 
 Agents first learn the best policy locally, and then at the end of the episode, a learner merges the Q-tables. 
 However, this does not reduce the problems of scalability and non-stationary, since a central agent must still collect all the Q-tables of the aggregate. 
 In particular, to handle the scalability issues, we will have to consider decentralised approaches based only on the neighbourhood information (or local leaders).

Finally, we do not address the multi-agent credit assignment problem in this work. 
 Indeed, we use a global reward function for the whole aggregate. In this way, however, learning often degenerated because it was difficult to understand the influence of an agent w.r.t. the system. 
 To reduce this problem, we will consider novel techniques of difference reward~\cite{DBLP:conf/atal/AgoginoT04} or counterfactual baseline~\cite{DBLP:journals/corr/FoersterFANW17}.
